---
title: "Reading Delimited Data"
format: html
toc: TRUE
editor_options: 
  chunk_output_type: inline
---

# Task 2

Packages

```{r}
library(readr)
library(dplyr)
```

# Glass Data

## Q1

Read in Data via url, delimiter is comma.

```{r}
glass_data <- read_csv("https://www4.stat.ncsu.edu/online/datasets/glass.data",
                        col_names= c("Id","RI","Na","Mg","Al","Si","K","Ca", "Ba", "Fe","Type_of_Glass"))

glass_data
```

## Q2 and Q3

Mutate **Type_of_Glass"** to go from numeric to descriptive via ifelse(). Only include observations wehre **Fe** \< 0.2 and **Type_of_Glass** is 'tableware' or 'headlamps'.

```{r}
glass_data |>
  mutate(Type_of_Glass= ifelse(Type_of_Glass==1,
                        "building_windows_float_processed",
                         ifelse(Type_of_Glass==2, 
                        "building_windows_non_float_processed",
                         ifelse(Type_of_Glass==3,
                        "vehicle_windows_float_processed",
                         ifelse(Type_of_Glass==4,
                         "vehicle_windows_non_float_processed",
                         ifelse(Type_of_Glass==5,
                         "containers",
                         ifelse(Type_of_Glass==6,
                          "tableware",
                         ifelse(Type_of_Glass==7,
                         "headlamps","other")))))))) |>
  filter(Fe<0.2,Type_of_Glass %in% c("tableware", "headlamps"))

```

# Yeast Data

## Q1

Read in data via url, double space delimited.

```{r}
yeast_data <- read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data",
                        col_names= c("seq_name","mcg","gvh","alm","mit","erl","pox","vac","nuc","class"))

yeast_data
```

## Q2 and Q3

Removing **seq_name** and **nuc** columns. Add columns for the mean and median of each numeric variable for each class grouping.

```{r}
yeast_data |>
  select(mcg:vac,class) |>
  group_by(class) |>
  mutate(across(where(is.numeric),list(mean = mean, median = median), .names = "{.col}_{.fn}"))
```
